---
title: "01_data_cleaning"
output: html_notebook
---


```{r}
library(tidyverse)
library(readxl)
```



```{r}
# load data
data_bsln <- read_csv("../data/data_bsln.csv")
data_plt <- read_csv("../data/data_plt.csv")

ditch_bsln <- read_csv("../data/ditch_bsln.csv")
ditch_plt <- read_csv("../data/ditch_plt.csv")
```

## Ditches

tidy up N/A in pilot
```{r}
ditch_plt <- ditch_plt %>% mutate(ditch_id = recode(ditch_id, "n/a" = NA_character_))
data_plt <- data_plt %>% mutate(ditch_id = recode(ditch_id, "n/a" = NA_character_))
```

ditches in pilot
```{r}
count(data_plt, ditch_id)
```


how many ditches numbered in pilot
```{r}
ditch_plt_comp <- ditch_plt %>% filter(ditch_id != is.na(ditch_id))
paste("there are", nrow(ditch_plt_comp), "numbered ditches in the pilot data")
ditch_plt_comp$ditch_id
```

```{r}
data_plt_comp <- data_plt %>% filter(ditch_id != is.na(ditch_id))
```


check ditch numbers in pilot exist in baseline
```{r}
ditch_plt_comp %>% mutate(in_baseline = ifelse(ditch_id %in% ditch_bsln$ditch_id, "1", "0")) %>% count(in_baseline)
```

## Species

how many species do we have

```{r}
spp_bsln <- unique(data_bsln$species)
spp_plt <- unique(data_plt$species)
spp_plt_comp <- unique(data_plt_comp$species)

paste("In the baseline there are", length(spp_bsln), "unique species names")
paste("In the pilot there are", length(spp_plt), "unique species names")
paste("In the pilot (numbered ditches) there are", length(spp_plt_comp), "unique species names")
```


How many species names in the new data match names in the old data:

```{r}
table(spp_plt %in% spp_bsln)
```
Jeepers...

lets just look at the numbered ditches: 
```{r}
table(spp_plt_comp %in% spp_bsln)
```

Jeepers still.  So lets look at these species:

```{r}
print("Species in both datasets:")
spp_plt_comp[spp_plt_comp %in% spp_bsln] %>% sort()
```

```{r}
print("Species in pilot not in baseline:")
spp_plt_comp[!spp_plt_comp %in% spp_bsln] %>% sort()
```

```{r}
print("Species in baseline not in pilot:")
spp_bsln[!spp_bsln %in% spp_plt_comp] %>% sort()
```

export to csv
```{r}
spp_bsln_t <- data.frame(species = spp_bsln, baseline = "Y")
spp_plt_t <- data.frame(species = spp_plt, pilot = "Y")
spp_plt_comp_t <- data.frame(species = spp_plt_comp, pilot_nbrd = "Y")
spp_comp <- spp_bsln_t %>% full_join(spp_plt_t) %>% full_join(spp_plt_comp_t) %>% arrange(species)
rm(spp_bsln_t, spp_plt_t, spp_plt_comp_t)
write_csv(spp_comp, "../data/species_comparison.csv")
```
## Match to GBIF

I did this in another script in which I also matched plabryatt to gbif. 

https://github.com/bogsnork/species_nbn_match 


## Match to NBN


```{r}
library(tidyverse)
```



```{r}
spp_comp <- read_csv("../data/species_comparison.csv"); spp_comp
```

## received cleaned species lookup from MC


```{r}
library(readxl)
spp_comp_annotated <- read_excel("../data/Species_names_MC annotations.xlsx") %>% 
  select(species_verbatim = species, 
         species_nbn = `current iRecord nomenclature`, 
         taxon_grp = 'taxonomic group',
         everything())
irecord_plt <- read_csv("../data/iRecord_download_MC_records.csv")
```



check overlap between pilot data and irecord data
```{r}
table(data_plt$gridref %in% irecord_plt$`Original map ref`)
```
all there

Write nbn codes to matched names

```{r}
tvks <- select(irecord_plt, Species, TVK = TaxonVersionKey) %>% distinct()
spp_lookup <- spp_comp_annotated %>% 
  left_join(tvks, by = c("species_nbn" = "Species"))
```

at this stage only the species present in pilot have TVKs.  In a way it doesn't matter because MC has checked for and found misspelled duplicates.  However, it may be an issue later.  



## Merge gbif lookups 
I did the lookup here: https://github.com/bogsnork/species_nbn_match 

```{r}
spp_comp_gbif <- read_csv("../data/species_comparison_gbif.csv")
```

```{r}
spp_lookup <- spp_lookup %>% 
  left_join(spp_comp_gbif, by=c("species_verbatim" = "species"))
```





Write matched names to data

```{r}
data_bsln_m <- data_bsln %>% 
  left_join(select(spp_lookup, species_verbatim, species_nbn, TVK, 
                   species_gbif = gbif.matched_name, gbif.taxon_id, 
                   taxon_grp), by = c("species" = "species_verbatim"))

data_plt_m <- data_plt %>% 
  left_join(select(spp_lookup, species_verbatim, species_nbn, TVK,
                   species_gbif = gbif.matched_name, gbif.taxon_id,  
                   taxon_grp), by = c("species" = "species_verbatim"))
```

subset data for analyis
```{r}
data_bsln_sub <- data_bsln_m %>% 
  filter(present != is.na(present)) %>% 
  filter(taxon_grp == "vascular plant") %>% 
  select(spp = species_nbn, ditch_id, abund = perc_freq, TVK,
         species_gbif, gbif.taxon_id,
         taxon_grp) %>% 
  mutate(ditch_id = as.character(ditch_id), 
         abund = as.character(abund))


data_plt_sub <- data_plt_m %>% 
  filter(ditch_id != is.na(ditch_id)) %>% 
  filter(taxon_grp == "vascular plant") %>% 
  select(spp = species_nbn, ditch_id, abund = plant_dafor, TVK,
         species_gbif, gbif.taxon_id, 
         taxon_grp) %>% 
  mutate(ditch_id = as.character(ditch_id), 
         abund = as.character(abund))

```

combind analysis data
```{r}
data_anls <- bind_rows(data_bsln_sub, data_plt_sub)
```






```{r}
write_csv(data_anls, "../data/data_analysis.csv")
write_csv(spp_lookup, "../data/spp_lookup.csv")
```





